\documentclass[12pt]{article}

\input{proposal-defs-alt}
\usepackage[margin=0.5in]{geometry}
\usepackage{url}
\usepackage{latexsym}
\usepackage{eepic,color,bm,array,amsmath}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{microtype}

\setlist[itemize]{leftmargin=1.25em, itemsep=0.05em, topsep=0.1em}
\setlength{\parskip}{0.15em}
\setlength{\parindent}{0pt}

\newcommand{\eg}{{\it e.g.}}
\newcommand{\ie}{{\it i.e.}}

\begin{document}

\pagestyle{empty}

\noindent
\begin{center}
{\large\bf Fairness-Aware Decision-Making in Quantum Routing and Clinical Settings\\ via Multi-Armed, Contextual, and Informative Bandit Methods}\\[0.2em]
{\large\textit{(Research Proposal)}}
\end{center}

\begin{center}
\begin{minipage}{0.48\linewidth}
\begin{center}
{\bf Piter Z. Garcia}\\
\begin{small}
{\footnotesize\mbox{MS Data Science / Decision-Making \& Algorithmic~Fairness}}\\
Rochester Institute of Technology (RIT)\\
{\it pizg8794@g.rit.edu}
\end{small}
\end{center}
\end{minipage}
\begin{minipage}{0.48\linewidth}
\begin{center}
{\bf Dr. Daniel Krutz, Travis Desell}\\
\begin{small}
Department of Software Engineering, Data Science\\
RIT Research Advisors\\
{\it dxkvse@g.rit.edu, tjdvse@g.rit.edu}
\end{small}
\end{center}
\end{minipage}
\end{center}

\noindent
% This project will develop a practical, reproducible framework for {\bf contextual multi-armed bandits} (iCMABs/\allowbreak{}CMABs)
% to support sequential decision-making under uncertainty, limited context, and non-stationarity while treating {\bf algorithmic fairness} as a first-class objective.
% The work on this generic framework focuses on measuring and improving fairness across the clinical (diagnostic-like) and quantum-network routing domains through the following deliverables:
% i) a simulation-first diagnostic-like sequential decision environment with controllable limited context, measurement noise, and distribution shift,
% ii) a fairness-aware contextual MAB evaluation stack (baselines + contextual policies) with time-evolving group disparity reporting and at least one mitigation mechanism,
% and iii) a transfer study demonstrating the same policy stack in a quantum-network routing simulator.

% Here, ``diagnostic-like'' refers to clinical diagnostic sequential decision workflows such as diagnostic test selection, triage or follow-up escalation, and model/pipeline selection under partial feedback and non-stationarity. It also includes cases where key diagnostic components (tests, instruments, or patient information) are unavailable or delayed due to resource constraints or allocation policies, creating group-dependent limited context.

% Mitigation mechanisms considered include fairness-regularized policy updates (constraints/penalties), group-aware calibration or thresholding, and missingness-aware context augmentation (e.g., feature acquisition or imputation policies).

This project develops a practical, reproducible framework for fairness-aware sequential
decision-making in two target settings: clinical diagnostic workflows and quantum-network
routing. The methodological focus is a spectrum of bandit methods ranging from
non-contextual to contextual to informative contextual bandits, designed to study how
different levels of available context affect fairness, equity, and performance under
uncertainty, non-stationarity, and resource constraints.

On the clinical side, the project focuses on sequential workflows such as diagnostic test
selection, triage, follow-up escalation, and model or pipeline choice when patient context
may be incomplete, delayed, noisy, or unevenly available across groups. On the quantum
side, the project focuses on routing and resource-allocation decisions under probabilistic link
success, scarce resources, and time-varying congestion, where service-equity gaps can
emerge across flow or user groups.

The initial deliverables are: i) a simulation-first clinical decision environment with
controllable missing context, measurement noise, and distribution shift; ii) a shared
evaluation stack for non-contextual, contextual, and informative bandit policies with
time-evolving disparity reporting, baseline unfairness assessment, and at least one
mitigation mechanism; and iii) a quantum-routing implementation of the same decision-making
framework to study fairness, robustness, and cross-domain transfer. Mitigation mechanisms
considered include fairness-regularized policy updates, group-aware calibration or
thresholding, and missingness-aware context augmentation.


% \paragraph{Background:} \hspace{-5mm}
% Many real-world workflows in both clinical diagnostics and quantum-network routing require sequential choices under uncertainty, resource constraints, and distribution shift. These settings can amplify inequities when some groups systematically have lower-quality context or different error profiles. This proposal frames these workflows as a contextual bandit problem: at each step, choose an action (an ``arm'') given observed context to maximize utility while controlling fairness gaps. In the quantum-network routing domain, routing decisions are constrained by probabilistic link success, limited entanglement/quantum-resource availability, and time-varying congestion; context may include link-quality estimates and network load/queue signals, but these signals can be delayed, noisy, or partially observed. In this project, service equity is defined as parity in routing success probability and latency across flow groups. In clinical diagnostics, context may include patient features, test/sample-quality indicators, and operational constraints, but access to context can be limited or systematically noisier for some populations. These limitations create performance and fairness risks, especially when optimizing aggregate performance, which can hide subgroup error spikes (e.g., false-negative gaps) unless the system is explicitly monitored and constrained. This project makes the time-evolving utility--fairness tradeoff explicit and tests transfer across both testbeds.


\paragraph{Background:} \hspace{-5mm}
Many real-world workflows in both clinical diagnostics and quantum-network routing require sequential choices under uncertainty, resource constraints, and distribution shift. These settings can amplify inequities when some groups systematically have lower-quality context or different error profiles. This project frames both as bandit-style decision problems: at each step, the system must choose an action given limited context, observe partial feedback, and balance utility with fairness over time.

In a clinical diagnostic workflow, the decision may involve which test to order next, whether to escalate follow-up, or which model or pipeline to trust when patient information is incomplete, delayed, or noisier for some groups. In that setting, fairness matters because optimizing only overall accuracy can hide subgroup failures, such as false-negative gaps or delayed care for populations whose context is systematically weaker.

In a quantum-network routing workflow, the decision may involve which path to route through or how to allocate scarce resources under probabilistic link success and time-varying congestion. In that setting, fairness appears as service equity: whether different flow or user groups receive comparable routing success, latency, and access to limited quantum resources rather than having one group consistently absorb the failures of the network.

By pairing these two settings, the proposal studies how different levels of context affect fairness, equity, and performance in sequential decision-making, while making the utility--fairness tradeoff explicit rather than treating it as a hidden side effect of optimization.


\paragraph{Related Work:} \hspace{-5mm}
This proposal builds on contextual bandit methods for sequential decision-making with side information \cite{li2010contextual,abbasi2011linear}. Algorithmic fairness work such as equality of opportunity \cite{hardt2016equality} motivates outcome-based disparity metrics, while fairness constraints in bandit learning emphasize the interaction between exploration and fairness requirements \cite{joseph2016fairness,chen2020fair}.


% \paragraph{Scientific Merit:} \hspace{-5mm}
% The core scientific question across the clinical diagnostics and quantum-network routing domains is: {\it when does informative context (and how it is modeled) materially reduce disparity and error in sequential decision-making under shift?} This is challenging in both domains because (1) feedback is partial and delayed (only the chosen test/routing action reveals an outcome), (2) conditions shift over time (patient mix and operational processes; network load and link conditions), and (3) fairness constraints can conflict with pure utility optimization. Practically, it builds on an existing quantum-network routing simulator and two prior works: threat-aware quantum routing using hybrid contextual bandits \cite{garcia2026threat_aware_quantum_routing} and {\it Equitable Bioinformatics} \cite{iste780_equitas}, operationalizing fairness audits and mitigation for diagnostic-relevant pipelines. The key innovative component is the two-domain evaluation: treating the {\bf quantum-network routing environment} as a first-class testbed alongside the diagnostic-like simulation, using the same fairness-aware contextual MAB stack in both to test robustness and transfer across radically different domains.

\paragraph{Scientific Merit:} \hspace{-5mm}
The core scientific question is: {\it how do different levels of decision context---from non-contextual to contextual to informative contextual bandit methods---change fairness, equity, and performance in sequential decision systems under shift?} This question is meaningful in both clinical diagnostics and quantum-network routing because each setting combines partial feedback, changing conditions, and resource constraints with the possibility that some groups systematically experience worse context or outcomes than others.

The challenge is not only to improve utility, but to understand when additional context materially reduces disparity, when it merely shifts who benefits, and when explicit fairness-aware mitigation is still required. This makes the problem scientifically interesting across both domains: clinical workflows expose high-stakes decision consequences for patient groups, while quantum routing exposes service-equity tradeoffs in resource-allocation systems under probabilistic failure and congestion.

Methodologically, the project builds on an existing quantum-network routing simulator and prior fairness-aware diagnostic work, but its main contribution is a shared evaluation framework for comparing non-contextual, contextual, and informative bandit methods across two very different settings while tracking utility--fairness tradeoffs over time. The scientific value of the two-domain design is not just transfer for its own sake, but testing whether the same fairness-aware decision principles remain meaningful across technically different but structurally similar sequential problems.



\paragraph{Broader Impacts:} \hspace{-5mm}
If successful, this project provides a concrete and reproducible way to study fairness-aware sequential decision-making in settings where limited context, scarce resources, and shifting conditions can produce unequal outcomes. In clinical diagnostic workflows, the work supports measuring and mitigating disparities that arise when some patient groups receive weaker context, delayed information, or lower-quality decision support. In quantum-network routing, it supports more trustworthy resource allocation by making service-equity gaps visible and testable rather than leaving them hidden inside aggregate routing performance.

More broadly, the project contributes a reusable evaluation framework for comparing bandit methods under fairness, equity, and performance objectives across more than one domain. That matters not only for this proposal's two testbeds, but also for the larger principle behind the work: fairness should be treated as part of the design and evaluation of sequential decision systems, not as a post-hoc audit added after optimization is already complete.


\paragraph{Approach:} \hspace{-5mm}
The goal is to implement and evaluate fairness-aware contextual MAB policies in two domains across two testbeds (diagnostic-like sequential decision simulation + quantum-network routing simulator), and to report utility--fairness tradeoffs under shift. The work is structured to be feasible without sensitive clinical data access (simulation-first), while remaining extensible to approved open datasets. The tasks are:
\begin{itemize}
\item {\bf Build diagnostic testbed:} implement a sequential decision simulation (test/model choice, retesting, pipeline choice) with distribution shift and controllable limited context and measurement noise.
\item {\bf Build quantum testbed:} integrate policies into the quantum-network routing simulator; define flow/node groups and track service equity across groups.
\item {\bf Implement bandit policies:} non-contextual baselines (epsilon-greedy, UCB, Thompson), a contextual baseline (LinUCB-style), and one informed contextual (iCMAB-style) policy under a shared interface.
\item {\bf Define reward + metrics:} utility with explicit cost/latency and safety weighting; report regret/utility plus time-evolving fairness metrics (diagnostics: group-wise FNR/FPR gaps; quantum routing: service-equity gaps in success probability and latency).
\item {\bf Add fairness mitigation:} implement one fairness-aware mechanism (constraint/penalty or calibration) and quantify utility--fairness tradeoffs across both testbeds.
\item {\bf Package for reproducibility:} modular scripts, fixed seeds + configs, and sanity checks for simulations and metric computation.
\end{itemize}
Phase 1 establishes baseline performance and group disparities under controlled limited context and measurement noise, evaluates an initial mitigation mechanism, and produces an initial demo; Phase 2 tests robustness under distribution shift and finalizes packaging for the final report/demo.


\paragraph{References:} \hspace{-5mm}
\scriptsize\begin{thebibliography}{10}
\bibitem{garcia2026threat_aware_quantum_routing}
P. Z. Garcia. {\it Threat-aware quantum routing via hybrid contextual bandits}. Internal manuscript (course/research draft), 2026.
\bibitem{iste780_equitas}
Garcia, P. Z. (2025). {\it Equitable Bioinformatics: Enhancing Diagnostic Decision-Making through RNA and Biomarker Data}. Unpublished paper, Rochester Institute of Technology.
\bibitem{li2010contextual}
L. Li, W. Chu, J. Langford, and R. Schapire. A contextual-bandit approach to personalized news article recommendation. In {\it WWW}, 2010.
\bibitem{abbasi2011linear}
Y. Abbasi-Yadkori, D. P{\'a}l, and C. Szepesv{\'a}ri. Improved algorithms for linear stochastic bandits. In {\it NeurIPS}, 2011.
\bibitem{hardt2016equality}
M. Hardt, E. Price, and N. Srebro. Equality of opportunity in supervised learning. In {\it NeurIPS}, 2016.
\bibitem{joseph2016fairness}
M. Joseph, M. Kearns, J. Morgenstern, and A. Roth. Fairness in learning: Classic and contextual bandits. In {\it NeurIPS}, 2016.
\bibitem{chen2020fair}
X. Chen, A. Zheng, Z. Zhou, and N. B. Shah. Fair contextual multi-armed bandits: Theory and experiments. In {\it UAI}, 2020.
\end{thebibliography}

\end{document}
