\section{Introduction}

\subsection{Scientific Merit, Importance, and Challenges}

\subsubsection{Scientific Merit}
Many applied data science systems are sequential decision systems (SDSs) that
route limited resources under uncertainty. SDSs are difficult because the
system observes only partial feedback. The chosen action reveals an outcome,
but the unchosen alternatives remain counterfactual.

\subsubsection{Importance}
SDSs are also fairness critical because the quality of the observed context is
often uneven across groups. A policy optimized only for aggregate reward can
hide subgroup errors and compound disparities over time when one group has
delayed testing, incomplete history, lower-quality measurements, or less
representative calibration data.

\subsubsection{Challenges}
This work frames that problem as fairness-aware contextual bandit learning. At
each round, a policy observes context, chooses an action, receives reward only
for the chosen action, and updates its future decisions. The scientific
questions focus not only on whether contextual information improves reward, but
on whether informative context and fairness-aware monitoring reduce disparities
when context is missing or degraded. This matters in clinical workflows because
false-negative or delayed-escalation gaps can produce direct harm. It also
matters in quantum-network routing because future communication systems will
allocate scarce probabilistic resources among competing flows; fairness in
success probability and latency is a service-quality issue.

\subsection{Overview of the Work and Relationship to Prior Work}

\subsubsection{Two Complementary Environment Families}
This work studies two complementary environment families. The first is a
quantum-routing evaluation stack based on hybrid contextual bandits for
entanglement routing and qubit allocation~\cite{garcia2026threat_aware_quantum_routing}.
That stack already exposes uncertainty, adversarial or non-stationary
conditions, legacy testbed conventions, and stateful evaluator artifacts. The
second is an embedded clinical simulation motivated by equitable bioinformatics
and fairness auditing work~\cite{garcia2025equitable_bioinformatics}.

\subsubsection{Clinical and Quantum Routing Analogy}
The clinical environment models the routing of scarce tests, retesting
capacity, escalation decisions, diagnostic attention, and model-assisted triage
across sites and cohorts. Under volatile conditions, such as a pandemic,
threat, demand, evidence quality, and resource availability can change
rapidly; this makes the clinical routing setting structurally comparable to
quantum routing under noisy links, uncertain entanglement generation, and
non-stationary service conditions. The two-domain design tests whether fairness
mechanisms and policy-update strategies transfer across fundamentally
different applications with the same partial-feedback allocation structure.

\subsubsection{Contribution Relative to Prior Work}
Across both environments, the central concern is that unequal context quality
can turn partial feedback into repeated allocation harm. The study therefore
evaluates performance and fairness as coupled outcomes rather than as separate
post-hoc measurements.

\subsection{Preliminary Results and Current Evidence}

\subsubsection{Selected MAB Spectrum}
Here, our MAB spectrum refers to the non-contextual, contextual, and
informative contextual bandit families evaluated in this work.

\subsubsection{Research Questions}
This study is organized around five research questions:

\begin{itemize}
    \item \textbf{RQ1:} How does context affect performance across our MAB spectrum?

    \item \textbf{RQ2:} What is the relationship between context, context quality, and fairness?

    \item \textbf{RQ3:} How does fairness-aware mediation affect overall performance?

    \item \textbf{RQ4:} Which MAB family is most likely to amplify disparities, and why?

    \item \textbf{RQ5:} In a non-stationary environment, which group is more likely to be harmed, and why?
\end{itemize}

\subsubsection{Current Evidence and Claim Boundary}
The current evidence supports a reproducible framework for studying these
questions across clinical and quantum routing settings. The quantum path
preserves legacy experiment compatibility while adding fairness artifacts. The
clinical path supports embedded evaluation across multiple model families,
allocation decisions, EQUITAS context, and model-stratified fairness summaries.
Claims about active fairness-adjusted quantum learning require an explicit
policy-update mechanism and full experimental validation.

\subsection{Research Overview}
The remainder of the paper is organized as follows: Section~II reviews related
work on bandits, fairness, missing context, clinical bias, and quantum routing;
Section~III presents the shared missing-context routing model and methodology;
Section~IV describes the implementation architecture; Section~V summarizes the
data and experimental inputs; Section~VI presents preliminary results;
Sections~VII and VIII discuss future work and conclusions.
