\begin{frame}{Break From Slides: GitHub Code Review}
\begin{block}{What to show live}
Open the GitHub repositories and show 2--3 classes that were not the focus of the midterm. Tie each class directly back to the architecture diagram.
\end{block}
\small
\begin{tabular}{p{0.30\textwidth}p{0.32\textwidth}p{0.25\textwidth}}
\toprule
Class / file & What to point out & Architecture layer \\
\midrule
\texttt{ClinicalExperimentEvaluator} & aggregation, fairness summaries, manifests & Layer 2 / Layer 4 \\
\texttt{ClinicalExperimentRunner} & seed/scenario execution and model loop & Layer 2 \\
\texttt{FairAllocatorV1} or \texttt{ClinicalBanditUCB} & allocation rule, update rule, context use & Layer 3 \\
\texttt{equitas\_mediator.py} & audit fields and mitigation context & Layer 4 \\
\bottomrule
\end{tabular}
\vspace{0.4em}
\begin{itemize}
\item Emphasize comments, formatting, and where fields/methods correspond to the architecture slides.
\item Explain low coupling: policies do not own reporting, and reporting does not own simulation.
\end{itemize}
\end{frame}

\begin{frame}{Working Demo Plan}
\begin{enumerate}
\item Run the clinical fairness smoke or demo notebook/script.
\item Show the generated fairness states and model-stratified summaries.
\item Open the manifest-linked report artifacts.
\item Show the COVID clinical foundation as evidence backing the clinical side.
\item Explain the quantum side as the retained legacy routing path plus next-semester active fairness updates.
\end{enumerate}
\vspace{0.4em}
\begin{block}{Demo goal for the rubric}
Demonstrate a working preliminary version of the project that executes the architecture path, records fairness evidence, and produces usable artifacts.
\end{block}
\end{frame}

\begin{frame}{What Gets Completed Next Semester}
\begin{enumerate}
\item Convert the COVID testing and pulse-ox cases into executable clinical environments.
\item Implement active iCMAB mitigation policies for missing or unreliable context.
\item Complete fairness-adjusted quantum policy updates, not only sidecar reporting.
\item Expand the validation hub into paper-ready efficiency and fairness reporting.
\item Produce final cross-domain experiments: reward/regret curves, disparity curves, and utility--fairness tradeoff plots.
\end{enumerate}
\end{frame}
