\documentclass{IEEEtran}

\usepackage{cite}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{url}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage[table]{xcolor}
\usepackage{tikz}
\usetikzlibrary{positioning,arrows.meta,fit,backgrounds}
\usepackage[hidelinks]{hyperref}

% Phase color palette
\definecolor{phase1blue}{RGB}{52,152,219}      % Environment / Input
\definecolor{phase2green}{RGB}{46,204,113}     % Context Observation
\definecolor{phase3orange}{RGB}{230,126,34}    % Decision / Action
\definecolor{phase4purple}{RGB}{155,89,182}    % Evaluation / Feedback

\title{Fairness-Aware Routing in Quantum Networks and Clinical Settings\\[0.3em]
{\large\textit{via Multi-Armed, Contextual, and Informative Bandit Methods}}\\[0.3em]
{\large\textit{(Approach Writeup)}}}

\author{%
\begin{minipage}{0.48\textwidth}
\begin{center}
{\bf Piter Z. Garcia}\\
\begin{small}
MS Data Science / Decision-Making \& Algorithmic Fairness\\
Rochester Institute of Technology (RIT)\\
{\it pizg8794@g.rit.edu}
\end{small}
\end{center}
\end{minipage}
\hfill
\begin{minipage}{0.48\textwidth}
\begin{center}
{\bf Dr. Daniel Krutz, Travis Desell}\\
\begin{small}
Department of Software Engineering, Data Science\\
RIT Research Advisors\\
{\it dxkvse@g.rit.edu, tjdvse@g.rit.edu}
\end{small}
\end{center}
\end{minipage}%
}

\begin{document}
\maketitle

%% ─────────────────────────────────────────────────────────────
\section{Approach Overview}
%% ─────────────────────────────────────────────────────────────

This work is grounded in a single unifying observation: both quantum-network
routing and clinical resource allocation are, at their core, \textit{routing
problems}. In quantum networking, entanglement and qubit resources must be
routed across competing paths and flows under probabilistic link success and
scarcity. In clinical settings, diagnostic and treatment resources---tests,
follow-ups, escalation decisions---must be routed among patients under
uncertainty, partial context, and operational constraints. In both cases,
\textit{unfair routing is not merely an equity problem; it is a performance
problem.} When one group of flows or patients is systematically deprioritized,
the downstream degradation in success probability, latency, or care quality
is direct and measurable. This shared structure makes the two domains
analytically equivalent and motivates a unified evaluation framework.

The methodological contribution is a shared, reproducible bandit-method
evaluation framework spanning three levels of context richness:
non-contextual multi-armed bandits (MABs), contextual multi-armed bandits
(CMABs), and informative contextual multi-armed bandits (iCMABs). The
framework is not merely applied in parallel to two domains; it uses the same
interaction model, fairness-logging pipeline, and comparison logic in both
settings, so that findings in one domain directly inform the other.

Two prior works anchor the testbeds. Equitable bioinformatics work provides
the clinical fairness foundation and subgroup disparity metrics
\cite{garcia2025equitable_bioinformatics}, while threat-aware quantum
routing work provides the domain-realistic network routing and resource-
allocation environment \cite{garcia2026threat_aware_quantum_routing}. The
cross-domain design is a deliberate research choice: quantum networking
represents an emerging frontier technology with proven potential to
transform communication, computation, and resource allocation. Pairing it
with a clinical setting that has immediate human impact---and sharing routing
as the structural bridge---positions this work at the intersection of present
necessity and future breakthrough.

%% ─────────────────────────────────────────────────────────────
\section{Motivating Workflows}
%% ─────────────────────────────────────────────────────────────

Two concrete routing workflows ground the shared domain model in
domain-specific operational constraints.

\subsection{Clinical Routing Workflow}
In the clinical setting, the decision-maker repeatedly routes resources:
ordering an initial screen, selecting a confirmatory test, or escalating
follow-up when a prior result is ambiguous. Each of these is a routing
decision that determines which patient receives which diagnostic or
therapeutic resource at which time. The observed context that drives these
decisions---patient features, prior test outcomes, sample quality indicators,
and operational constraints---may be incomplete, delayed, or noisier for
some demographic groups. When context quality is uneven, the routing policy
may systematically under-resource certain subgroups, producing higher
false-negative rates, delayed escalation, or inferior downstream care.
Clinical fairness is therefore treated as a time-evolving property of the
routing process rather than a static post-hoc audit of model outputs.

\subsection{Quantum Routing Workflow}
In the quantum-network setting, the decision-maker routes entanglement:
selecting paths and allocating scarce entanglement-generation attempts among
competing flows under probabilistic link success and changing network
conditions. Observed routing context---link-quality estimates, queue and load
signals, or threat indicators---can be noisy, stale, or only partially
observed. When context is unreliable, routing policies may consistently
assign degraded paths or insufficient resources to certain flow groups,
creating service-equity gaps in success probability, latency, and resource
access. This interpretation makes quantum routing a genuine fairness domain
rather than a purely performance-oriented benchmark.

%% ─────────────────────────────────────────────────────────────
\section{Shared Domain Model}
%% ─────────────────────────────────────────────────────────────

The shared domain model makes the common routing-decision structure explicit
across both testbeds while preserving the domain-specific interpretation of
states, actions, rewards, and fairness metrics. The model is organized into
four phases, each color-coded in Figure~\ref{fig:domain_model}: (1)
\textcolor{phase1blue}{\textbf{Environment / Input}} (blue), (2)
\textcolor{phase2green}{\textbf{Context Observation}} (green), (3)
\textcolor{phase3orange}{\textbf{Routing Decision}} (orange), and (4)
\textcolor{phase4purple}{\textbf{Evaluation and Mitigation}} (purple).

\begin{figure*}[!t]
\centering
\resizebox{0.97\textwidth}{!}{%
\begin{tikzpicture}[
  font=\scriptsize,
  box/.style={draw, rounded corners=4pt, align=center,
              text width=1.75cm, minimum height=0.78cm,
              inner sep=3pt, text=white, font=\scriptsize\bfseries},
  phase1/.style={box, fill=phase1blue},
  phase2/.style={box, fill=phase2green},
  phase3/.style={box, fill=phase3orange},
  phase4/.style={box, fill=phase4purple},
  arr/.style={-Stealth, thick, gray!60},
  label/.style={font=\scriptsize\bfseries, align=center}
]

%% Phase 1 — Environment / Input (blue)
\node[phase1] (clin)  {Clinical\\Routing};
\node[phase1, below=0.5cm of clin]  (quant) {Quantum\\Routing};
\node[phase1, right=0.6cm of clin, yshift=-0.9cm] (env)  {Environment /\\Testbed};

%% Phase 2 — Context Observation (green)
\node[phase2, right=0.6cm of env]  (ctx) {Observed\\Context};
\node[phase2, above=0.55cm of ctx] (deg) {Missing / Noisy /\\Delayed Context};

%% Phase 3 — Routing Decision (orange)
\node[phase3, right=0.6cm of ctx]  (pol) {Policy Family\\MAB / CMAB / iCMAB};
\node[phase3, right=0.6cm of pol]  (act) {Routing\\Action};

%% Phase 4 — Evaluation / Feedback (purple)
\node[phase4, below=0.55cm of act] (fb)  {Partial Feedback\\and Reward};
\node[phase4, right=0.6cm of act]  (log) {Utility + Fairness\\Logger};
\node[phase4, below=0.55cm of log] (mit) {Mitigation +\\Comparison Loop};

%% Arrows
\draw[arr] (clin)  -- (env);
\draw[arr] (quant) -- (env);
\draw[arr] (env)   -- (ctx);
\draw[arr] (deg)   -- (ctx);
\draw[arr] (ctx)   -- (pol);
\draw[arr] (pol)   -- (act);
\draw[arr] (act)   -- (fb);
\draw[arr] (act)   -- (log);
\draw[arr] (fb)    -- (log);
\draw[arr] (log)   -- (mit);

%% Phase background bands
\begin{pgfonlayer}{background}
  \node[fill=phase1blue!15, rounded corners=6pt, inner sep=4pt,
        fit=(clin)(quant)(env)] (bg1) {};
  \node[fill=phase2green!15, rounded corners=6pt, inner sep=4pt,
        fit=(ctx)(deg)] (bg2) {};
  \node[fill=phase3orange!15, rounded corners=6pt, inner sep=4pt,
        fit=(pol)(act)] (bg3) {};
  \node[fill=phase4purple!15, rounded corners=6pt, inner sep=4pt,
        fit=(fb)(log)(mit)] (bg4) {};
\end{pgfonlayer}

%% Phase labels below each band
\node[label, text=phase1blue, below=0.12cm of bg1] {\textbf{Phase 1}\\Environment};
\node[label, text=phase2green, below=0.12cm of bg2] {\textbf{Phase 2}\\Context};
\node[label, text=phase3orange, below=0.12cm of bg3] {\textbf{Phase 3}\\Decision};
\node[label, text=phase4purple, below=0.12cm of bg4] {\textbf{Phase 4}\\Evaluation};

\end{tikzpicture}%
}
\caption{%
  Shared domain model across both testbeds, organized into four
  color-coded phases.
  \textcolor{phase1blue}{\textbf{Phase 1 (blue) --- Environment / Input:}}
    Both the clinical routing workflow and the quantum routing workflow feed
    into a shared testbed interface.
  \textcolor{phase2green}{\textbf{Phase 2 (green) --- Context Observation:}}
    The system observes available routing context, subject to missingness,
    noise, or delay---an explicit part of the research question, not an
    implementation nuisance.
  \textcolor{phase3orange}{\textbf{Phase 3 (orange) --- Routing Decision:}}
    A policy from the selected method family (MAB, CMAB, or iCMAB) selects
    a routing action under the observed context.
  \textcolor{phase4purple}{\textbf{Phase 4 (purple) --- Evaluation and Mitigation:}}
    The environment returns partial feedback; utility and fairness metrics
    are logged per step; and baseline and fairness-aware mitigation policies
    are compared within the same loop.}
\label{fig:domain_model}
\end{figure*}

The four-phase structure is identical in both domains. What differs is the
operational semantics: in the clinical domain, a routing action is a
diagnostic or treatment decision; in the quantum domain, it is a path
selection or entanglement-allocation decision. Fairness in both cases is
measured as a disparity in outcomes across subgroups---patient demographics
in the clinical setting, flow or user groups in the quantum setting.

%% ─────────────────────────────────────────────────────────────
\section{Design Rationale and Novelty}
%% ─────────────────────────────────────────────────────────────

\subsection{Why this approach has not been done before}
No prior work evaluates fairness-aware sequential routing under a unified
bandit framework that spans both clinical resource allocation and quantum
network entanglement routing. Existing literature addresses these domains in
isolation: contextual bandits for clinical trials focus on treatment
optimization without modeling routing fairness across patient subgroups
\cite{garcia2025equitable_bioinformatics}; quantum routing research
optimizes throughput and fidelity without examining service-equity
disparities under partial and noisy context. The critical gap is the
recognition that routing unfairness and routing performance are inseparable
in both domains. Addressing this gap requires a shared evaluation framework
that can detect, compare, and mitigate routing disparities under the same
experimental conditions---something no existing approach provides.

Furthermore, quantum networking is not a distant hypothetical. It is a
proven, emerging technology with documented potential to reshape
communication, computation, and resource distribution. Pairing it with a
clinical domain that directly affects human health and equity is not an
arbitrary coupling; it is a deliberate positioning of this research at the
frontier where technological advancement and human fairness intersect.

\subsection{Why this approach is better than others}
Single-domain approaches cannot test whether a fairness principle is
structurally general or merely domain-specific. A study limited to clinical
bandits or to quantum routing alone provides no evidence that the findings
transfer or that the framework is robust across different state spaces,
reward structures, and fairness definitions. By sharing the same interaction
model, logging pipeline, and comparison logic across two structurally
analogous but operationally different routing domains, this approach produces
findings that are simultaneously more credible and more generalizable.

Approaches that apply fairness as a post-hoc audit---measuring disparity
after training or deployment---cannot capture how routing decisions compound
over time and how fairness evolves under learning. This framework measures
fairness at every step of the decision process, making it possible to detect
early-onset disparities, evaluate mitigation mechanisms in real time, and
compare policies under the same partial-feedback conditions that make routing
decisions difficult in practice.

\subsection{Design and implementation decisions}
\textit{Bandit methods over full reinforcement learning.} The bandit
formulation is chosen because both routing domains involve partial feedback:
the decision-maker observes only the outcome of the chosen route, not the
counterfactual outcomes of unchosen alternatives. Full RL would require
richer state-transition modeling that is neither available nor necessary for
the fairness questions being posed.

\textit{Simulation-first clinical testbed.} A simulation environment for
clinical routing is used intentionally so that missingness, noise, and
group-specific context degradation can be introduced under controlled,
reproducible conditions without dependence on sensitive patient data. This
is not a limitation---it is the correct choice for isolating the effect of
routing context quality on fairness outcomes.

\textit{Extension of existing quantum routing environment.} Rather than
inventing a second quantum domain, the approach extends an already-grounded,
threat-aware quantum routing environment
\cite{garcia2026threat_aware_quantum_routing}. This decision preserves
domain realism, avoids the credibility risk of a fabricated testbed, and
ensures that quantum routing findings reflect conditions studied in prior
work.

\textit{Three-level context spectrum (MAB, CMAB, iCMAB).} The spectrum is
not a heuristic choice. It directly operationalizes the core research
question: whether routing fairness can be improved through richer context, or
whether context itself, when unevenly available, amplifies rather than
reduces disparities. The three levels create a controlled ladder of
information richness that isolates the effect of context quality.

\subsection{How these decisions relate to the domain model}
Every design decision maps directly to a phase in the shared domain model
(Figure~\ref{fig:domain_model}). The choice of bandit methods over full RL
is a Phase~3 decision: it constrains the policy family to methods that
operate under partial feedback, which is also the feedback structure of
Phase~4. The simulation-first clinical design and the quantum environment
extension are Phase~1 decisions: they determine what inputs enter the
testbed and under what conditions. The three-level context spectrum is a
Phase~2 decision: it controls what context is observable and how degraded
that context can be. The shared fairness-logging and mitigation loop is a
Phase~4 decision: it ensures that fairness is a first-class, per-step
output rather than an afterthought. The domain model is therefore not a
diagram added after the fact; it is the organizational structure that
determined each design choice.

%% ─────────────────────────────────────────────────────────────
\section{Implementation and Evaluation Plan}
%% ─────────────────────────────────────────────────────────────

The implementation plan prioritizes reproducibility and systematic fairness
evaluation across both testbeds. The clinical simulation environment models
sequential routing decisions---test ordering, retesting, and escalation---
under conditions where some patient groups have weaker or noisier observed
context. The quantum routing environment represents path selection and
entanglement-resource allocation under probabilistic link success, changing
load, and limited quantum resources. Both testbeds expose the same policy
interface so that MAB, CMAB, and iCMAB methods are evaluated under a shared
logging and analysis pipeline.

The reward layer tracks utility-oriented routing outcomes: accuracy, cost,
success rate, and latency. The fairness layer tracks domain-specific routing
disparities: clinical false-negative and delayed-escalation gaps across
patient subgroups; quantum service-equity gaps in routing success or latency
across flow groups. Fairness is therefore a first-class output of the
learning process.

Experiments run from explicit configuration files with fixed random seeds and
per-step metric logging. At least one fairness-aware mitigation mechanism is
evaluated inside the same framework so that the final comparison includes
both baseline disparity assessment and the effect of active mitigation under
learning.

%% ─────────────────────────────────────────────────────────────
\section{Research Contribution}
%% ─────────────────────────────────────────────────────────────

This work makes three distinct contributions. First, it identifies routing
as the structural common ground between clinical resource allocation and
quantum network entanglement distribution, and builds the first fairness-
aware sequential decision framework that evaluates both under identical
experimental conditions.

Second, it provides the first systematic comparison of non-contextual,
contextual, and informative-context bandit policies specifically with respect
to routing fairness, making it possible to determine whether richer context
reduces routing disparities or introduces new ones when context quality is
uneven across groups.

Third, it advances a new direction in applied fairness research by connecting
a domain with immediate human health impact---clinical routing---to an
emerging breakthrough technology---quantum networking---through a shared
structural property. This positions the findings not only as an evaluation
of current systems, but as a foundation for designing fairer routing
mechanisms in next-generation quantum-enabled infrastructures.

\begin{thebibliography}{9}\setlength{\itemsep}{0em}
\bibitem{garcia2025equitable_bioinformatics}
Garcia, P. Z. Equitable Bioinformatics: Enhancing Diagnostic
Decision-Making through {RNA} and Biomarker Data. Course paper
(ISTE~780 Phase~4), Rochester Institute of Technology, 2025.
\bibitem{garcia2026threat_aware_quantum_routing}
Garcia, P. Z. Threat-Aware Evaluation of Quantum Entanglement Routing
and Qubit Allocation via Hybrid Contextual Bandits. Unpublished
manuscript (GA-Work internal draft), Rochester Institute of Technology,
2026.
\end{thebibliography}

\end{document}