\documentclass{IEEEtran}

\usepackage{cite}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{url}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage[table]{xcolor}
\usepackage{tikz}
\usetikzlibrary{positioning,arrows.meta,fit,backgrounds,calc}
\usepackage[hidelinks]{hyperref}

\definecolor{DeepNavy}{RGB}{11,26,51}
\definecolor{SoftGray}{RGB}{245,247,250}

\definecolor{phase1blue}{RGB}{52,152,219}
\definecolor{phase2green}{RGB}{46,204,113}
\definecolor{phase3orange}{RGB}{230,126,34}
\definecolor{phase4purple}{RGB}{155,89,182}

\title{{\LARGE\textbf{Fairness-Aware Routing in Quantum Networks \& Clinical Settings}}\\
{\Large\textit{via Multi-Armed, Contextual, and Informative Bandit Methods}}\\
{\large\texttt{(Approach Writeup)}}}

\author{%
\begin{minipage}{0.48\textwidth}
\begin{center}
{\bf Piter Z. Garcia}\\
\begin{small}
MS Data Science / Decision-Making \& Algorithmic Fairness\\
Rochester Institute of Technology (RIT)\\
{\it pizg8794@g.rit.edu}
\end{small}
\end{center}
\end{minipage}
\hfill
\begin{minipage}{0.48\textwidth}
\begin{center}
{\bf Dr. Daniel Krutz, Travis Desell}\\
\begin{small}
Department of Software Engineering, Data Science\\
RIT Research Advisors\\
{\it dxkvse@g.rit.edu, tjdvse@g.rit.edu}
\end{small}
\end{center}
\end{minipage}%
}

\begin{document}
\maketitle

%% ─────────────────────────────────────────────────────────────
\section{Approach Overview}
%% ─────────────────────────────────────────────────────────────

Quantum-network routing and clinical resource allocation are, at their
core, \textit{routing problems}---and both are vulnerable to the same failure
mode: \textit{missing context}. In quantum networking, entanglement and qubit
resources must be routed across competing flows under probabilistic link
success and scarcity; standard policies can underperform because they ignore,
misuse, or lose the context that would guide fair allocation. In clinical
settings, diagnostic and treatment resources are routed among patients under
uncertainty and partial information; tools calibrated on non-representative
populations can become less responsive to the context of those they exclude.
In both cases, the groups whose context is missing or weakly represented can
absorb the worst outcomes---and the degradation is a direct performance failure. 
This shared root cause makes the two domains comparable and 
motivates a unified evaluation work.

The contribution is a reproducible bandit-method framework spanning three
context-richness levels---non-contextual MABs, contextual CMABs, and
informative iCMABs---designed specifically to probe what happens when context
is absent, partially available, or actively recovered. Two prior works anchor
the testbeds: equitable bioinformatics work provides the clinical fairness
foundation \cite{garcia2025equitable_bioinformatics}, and threat-aware
quantum routing work provides the network environment
\cite{garcia2026threat_aware_quantum_routing}. Quantum networking is a rapidly
advancing technology likely to influence future communication and resource
allocation; pairing it with clinical fairness through the structural bridge
of missing-context routing positions this work between an urgent present-day
problem and an important next-generation systems challenge. The overall
structure is summarized in Figure~\ref{fig:domain_model}(a), and the shared
execution harness used to run both testbeds is detailed in
Figure~\ref{fig:domain_model}(b).

%% ─────────────────────────────────────────────────────────────
\section{Motivating Workflows}
%% ─────────────────────────────────────────────────────────────

\subsection{Clinical Routing Workflow}
The clinical decision-maker routes resources---screens, confirmatory tests,
escalation follow-ups---among patients under incomplete, delayed, or
group-unequal context. During the COVID-19 response, uneven access to testing,
follow-up, and related resources made the consequences of poor routing
especially visible, contributing to delayed detection and weaker care pathways
for some communities. A compounding failure reveals a second dimension:
diagnostic protocols built from non-representative control groups calibrated
thresholds against an incomplete population, making certain groups less visible. The
result is missing context at the point of decision, which can amplify risk for
the very groups bearing the highest burden.

\subsection{Quantum Routing Workflow}
The quantum decision-maker routes entanglement by selecting paths and
allocating scarce entanglement-generation attempts among competing flows
under noisy, stale, or partial context. Standard MABs ignore context
entirely; CMABs use it but remain vulnerable to noise and missingness.
Empirical testing suggests that iCMABs can outperform both---because they
recover and correctly weight context the other methods miss or misuse.
This is the same structural problem. Missing context
in the routing policy produces the same outcome pattern seen in the clinical
instrument: the groups whose information is least represented tend to absorb
the worst outcomes.

%% ─────────────────────────────────────────────────────────────
\section{Shared Domain Model}
%% ─────────────────────────────────────────────────────────────

The domain model in Figure~\ref{fig:domain_model}(a) organizes the
routing-decision process into four phases---
\textcolor{phase1blue}{\textbf{Environment}} (blue),
\textcolor{phase2green}{\textbf{Context Observation}} (green),
\textcolor{phase3orange}{\textbf{Routing Decision}} (orange),
\textcolor{phase4purple}{\textbf{Evaluation and Mitigation}} (purple), structurally 
consistent across testbeds. Operational semantics differ: a routing action is 
a diagnostic decision in the clinical domain and a path-selection decision in the 
quantum domain; fairness is measured as outcome disparity across patient or network flow groups.

The COVID-19 response offers a practical illustration of what this domain model is
designed to study. Diagnostic tests were scarce, unevenly distributed, and
often delayed, making clear that unfair routing can become a performance
failure rather than only an abstract equity concern.

What makes this framework especially useful is that both domains also share the same core failure mode: missing
context. In clinical settings, excluded populations can become less visible to
the instrument. In quantum routing, iCMABs can outperform MABs and CMABs not because of complexity, but because they recover context the
other methods discard or underuse. The underperforming groups in both cases
are often those whose context was never properly captured.
\textit{Phase~2 is therefore where the root cause of routing unfairness often
enters the system}---and the primary phase this framework is designed to interrogate.
\textit{Phase~4}, shown conceptually in Figure~\ref{fig:domain_model}(a) and
operationally expanded in Figure~\ref{fig:domain_model}(b), then detects,
logs, and mitigates the downstream consequences before they compound.

%% ─────────────────────────────────────────────────────────────
\section{Design Rationale and Novelty}
%% ─────────────────────────────────────────────────────────────

\subsection{Why This Approach}
To our knowledge, prior work has not evaluated fairness-aware sequential
routing under one unified bandit framework spanning clinical and quantum
domains, and prior work has not centered missing context as the shared
structural driver of routing unfairness across both. Clinical bandit research
often focuses on treatment optimization without examining how missing
population context degrades routing equity; quantum routing research often
focuses on throughput, reliability, or threat response without directly
connecting context missingness to service-equity disparities. Studying them
together creates findings that can improve today's clinical systems while
informing fair routing in tomorrow's quantum-enabled infrastructures.

\subsection{Why This Approach Is Stronger Than Simpler Alternatives}
Single-domain approaches cannot verify whether missing context drives routing
unfairness structurally or only situationally; findings remain confined to
one state space and one fairness definition. Post-hoc audits cannot capture
how context gaps compound into routing disparities over time under learning.
This framework measures fairness at every step under the same partial-feedback
conditions that make routing difficult in practice, enabling real-time
detection of where/when missing context causes disparity, and direct
comparison among policies that ignore, partially use, or actively recover
that context.

\subsection{Design and Implementation Decisions}
\textit{Bandits over full RL:} both domains return only partial feedback,
making bandit methods the correct formulation. \textit{Simulation-first
clinical testbed:} it allows controlled injection of missing context,
noise, and group-specific degradation without sensitive patient data.
\textit{Existing quantum environment extended:} reusing a grounded,
threat-aware routing environment preserves domain-realistic missing-context
conditions empirically associated with iCMAB performance.
\textit{MAB / CMAB / iCMAB spectrum:} each directly operationalizes the core
question---\textit{does recovering missing routing context reduce disparities, or are
additional fairness mechanisms still required?}

\subsection{Connecting Decisions to the Domain Model}
Each design decision corresponds to a deliberate choice about what a specific
phase in Figure~\ref{fig:domain_model}(a) must do and why, while the shared
pipeline in Figure~\ref{fig:domain_model}(b) shows how those choices are
executed consistently across both testbeds.

\textsc{Phase~1 (Environment)} is shaped by the simulation-first clinical
testbed and the quantum environment extension. The simulation introduces
missing and group-unequal context as a controlled variable, making
\textsc{Phase~1} a research instrument, not just a data source. The quantum
extension means \textsc{Phase~1} inherits domain-realistic conditions---link success
rates, entanglement scarcity, threat signals---that reproduce the same
missing-context failure mode without fabricating a testbed.

\textsc{Phase~2 (Context Observation)} is shaped by the MAB/CMAB/iCMAB
spectrum. Each level defines how much context reaches the policy: none,
structured but potentially incomplete, or actively recovered despite noise
and missingness. \textsc{Phase~2} is the experimental variable and the locus of the
root cause; spanning three levels helps measure how much of the routing disparity is
explained by missing context alone.

\textsc{Phase~3 (Routing Decision)} is shaped by the bandit choice. Because
both environments return only the outcome of the chosen route---never the
counterfactual---the policy must be a bandit policy. Full RL would require
transition-model information that does not exist. The bandit formulation is
the most structurally honest choice given what \textit{Phase~4} can return.

\textsc{Phase~4 (Evaluation and Mitigation)} justifies the entire design.
Logging utility and fairness at every step produces a per-step disparity
trajectory, making it possible to pinpoint exactly when missing context
causes disparities to emerge, persist, or resolve under mitigation. The
domain model is the organizational structure that determined every design
choice from \textit{Phase~1} through \textsc{Phase~4}.

%% ─────────────────────────────────────────────────────────────
\section{Implementation and Evaluation Plan}
%% ─────────────────────────────────────────────────────────────

Both testbeds expose the same policy interface, so MAB, CMAB, and iCMAB
methods are evaluated under the shared logging and analysis pipeline shown in
Figure~\ref{fig:domain_model}(b). The
clinical simulation models test ordering, retesting, and escalation under
deliberately introduced missing and group-unequal context; the quantum
environment reproduces domain-realistic context missingness that aligns with
the empirically observed iCMAB advantage. The reward layer tracks accuracy,
cost, success rate, and latency; the fairness layer tracks clinical
false-negative and delayed-escalation gaps and quantum service-equity gaps
across flow groups. Experiments use explicit configurations, fixed
random seeds, and per-step metric logging; at least one fairness
mitigation mechanism is evaluated to compare policies that
ignore, partially use, or actively recover missing context.

%% ─────────────────────────────────────────────────────────────
\section{Research Contribution}
%% ─────────────────────────────────────────────────────────────

This work makes three contributions. First, it identifies routing and missing
context as the shared structural ground between clinical and quantum domains,
and builds a unified fairness-aware sequential decision framework for
evaluating both under comparable conditions. Second, it provides a systematic
MAB/CMAB/iCMAB comparison with respect to routing fairness, directly
measuring how much disparity is explained by missing context and how much
requires additional mitigation. Third, by connecting a domain of immediate
human health impact to an important emerging technology through their shared
missing-context failure mode, the findings serve both as an evaluation of
current systems and a foundation for fair routing in next-generation
quantum-enabled infrastructures.

%% ─── Figure at end — does not count toward two-page limit ───

\begin{thebibliography}{9}\setlength{\itemsep}{0em}
\bibitem{garcia2025equitable_bioinformatics}
Garcia, P. Z. Equitable Bioinformatics: Enhancing Diagnostic
Decision-Making through {RNA} and Biomarker Data. Course paper
(ISTE~780 Phase~4), Rochester Institute of Technology, 2025.
\bibitem{garcia2026threat_aware_quantum_routing}
Garcia, P. Z. Threat-Aware Evaluation of Quantum Entanglement Routing
and Qubit Allocation via Hybrid Contextual Bandits. Unpublished
manuscript, Rochester Institute of Technology, 2026.
\end{thebibliography}

% Drop-in replacement for Figure 1 in the approach writeup.
% Requires: xcolor, tikz
% TikZ libraries: positioning, arrows.meta, fit, backgrounds, calc
\begin{figure*}[!t]
\centering
\resizebox{0.98\textwidth}{!}{%
\begin{tikzpicture}[
  font=\scriptsize,
  hi/.style={
    draw, rounded corners=4pt, align=center, text width=1.78cm,
    minimum height=0.80cm, inner sep=3pt, text=white,
    font=\scriptsize\bfseries
  },
  phase1/.style={hi, fill=phase1blue},
  phase2/.style={hi, fill=phase2green},
  phase3/.style={hi, fill=phase3orange},
  phase4/.style={hi, fill=phase4purple},
  pbox/.style={rounded corners=7pt, inner sep=7pt, line width=0.9pt},
  arr/.style={-Stealth, line width=1.0pt, draw=black!80},
  zoomarr/.style={-Stealth, dashed, line width=1.0pt, draw=phase4purple!85!black},
  label/.style={font=\scriptsize\bfseries, text=black!85},
  smallnote/.style={font=\scriptsize, text=black!70, align=center},
  zoomtag/.style={
    draw=phase4purple!80!black, fill=white, rounded corners=4pt,
    inner sep=3pt, font=\scriptsize\bfseries, text=phase4purple!85!black
  }
]

% =========================================================
% Figure 1(a): high-level shared model
% =========================================================
\node[label, anchor=west] (a_title) at (-0.15,2.50)
{Figure 1(a). Shared domain model};

\node[phase1] (clin) at (0.8,1.35) {Clinical\\Routing};
\node[phase1, below=0.55cm of clin] (quant) {Quantum\\Routing};
\node[phase1, right=0.75cm of clin, yshift=-0.92cm] (env)
{Environment /\\Testbed};

\node[phase2, right=0.80cm of env] (ctx) {Observed\\Context};
\node[phase2, above=0.58cm of ctx] (deg) {Missing / Noisy /\\Delayed Context};

\node[phase3, right=0.80cm of ctx] (pol) {Policy Family\\MAB / CMAB / iCMAB};
\node[phase3, right=0.80cm of pol] (act) {Routing\\Action};

\node[phase4, right=0.85cm of act, text width=2.18cm, minimum height=1.55cm] (pipe)
{Shared Evaluation\\Pipeline\\\scriptsize see Figure 1(b)};

\draw[arr] (clin.east) -- (env.west);
\draw[arr] (quant.east) -- (env.west);
\draw[arr] (env) -- (ctx);
\draw[arr] (deg) -- (ctx);
\draw[arr] (ctx) -- (pol);
\draw[arr] (pol) -- (act);
\draw[arr] (act) -- (pipe);

\begin{scope}[on background layer]
  \node[pbox, fill=phase1blue!15, draw=phase1blue!80!black, fit=(clin)(quant)(env)] (bg1) {};
  \node[pbox, fill=phase2green!15, draw=phase2green!80!black, fit=(ctx)(deg)] (bg2) {};
  \node[pbox, fill=phase3orange!15, draw=phase3orange!80!black, fit=(pol)(act)] (bg3) {};
  \node[pbox, fill=phase4purple!15, draw=phase4purple!80!black, fit=(pipe)] (bg4) {};
\end{scope}

\node[label, text=phase1blue!80!black, below=0.16cm of bg1] {Phase 1: Environment};
\node[label, text=phase2green!70!black, below=0.16cm of bg2] {Phase 2: Context};
\node[label, text=phase3orange!80!black, below=0.16cm of bg3] {Phase 3: Decision};
\node[label, text=phase4purple!80!black, below=0.16cm of bg4] {Phase 4: Pipeline module};

% =========================================================
% Figure 1(b): horizontal zoom-in of Phase 4
% =========================================================
\node[rounded corners=9pt, draw=phase4purple!80!black, fill=phase4purple!15,
      minimum width=16.3cm, minimum height=3.10cm, anchor=north west] (panel)
      at (-0.15,-1.80) {};

\node[label, anchor=west] at ($(panel.north west)+(0.18,-0.28)$)
{Figure 1(b). Zoom-in of Figure 1(a), Phase 4: Shared evaluation pipeline};

\node[zoomtag, anchor=east] at ($(panel.north east)+(-0.18,-0.28)$)
{Submodule of Figure 1(a)};

\node[phase4, text width=1.95cm, minimum height=0.95cm, anchor=west] (cfg)
at (0.15,-2.95) {Config\\\scriptsize missingness + shift};

\node[phase4, text width=2.00cm, minimum height=0.95cm, right=0.36cm of cfg] (tb)
{Testbed\\\scriptsize clinical or quantum};

\node[phase4, text width=1.95cm, minimum height=0.95cm, right=0.36cm of tb] (pp)
{Policy\\\scriptsize MAB / CMAB / iCMAB};

\node[phase4, text width=1.95cm, minimum height=0.95cm, right=0.36cm of pp] (lg)
{Logger\\\scriptsize full trajectories};

\node[phase4, text width=2.05cm, minimum height=0.95cm, right=0.36cm of lg] (mt)
{Metrics\\\scriptsize utility + disparity($t$)};

\node[phase4, text width=2.05cm, minimum height=0.95cm, right=0.36cm of mt] (mg)
{Mitigation\\\scriptsize constraints + ablations};

\draw[arr] (cfg) -- (tb);
\draw[arr] (tb) -- (pp);
\draw[arr] (pp) -- (lg);
\draw[arr] (lg) -- (mt);
\draw[arr] (mt) -- (mg);

\node[smallnote, anchor=south] at ($(panel.south)+(0,0.18)$)
{One shared harness across domains: configure, run policies, log full state, compute fairness and utility, then compare or mitigate.};

\draw[zoomarr] (bg4.south) to[out=-90,in=90] ($(panel.north)+(4.70,0.02)$);

\end{tikzpicture}%
}
\caption{
Shared domain model with an embedded zoom-in.
\textbf{Figure 1(a)} shows the four high-level modules of the cross-domain routing framework.
\textbf{Figure 1(b)} is a zoom-in of \textbf{Phase 4} from Figure 1(a), showing the shared evaluation pipeline used across both testbeds: configuration, testbed selection, policy execution, logging, metrics, and mitigation.
The purple background in Figure 1(b) matches the purple pipeline module in Figure 1(a) to make the zoom relationship explicit.
}
\label{fig:domain_model}
\end{figure*}

\end{document}